Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-09-23 bbbq_article.tex
Process feedback @bogeert, progress https://github.com/richelbilderbeek/bbbq_article/issues/69, move text about HLA classification to https://github.com/richelbilderbeek/bbbq_article/issues/110

The human immune system is unexpectedly attentive towards
human transmembrane helices (TMHs), as TMH-derived peptide fragments
bind to histocompatibility complex (MHC) class I more often then expected
by chance only.
This finding hints that there has been selection upon detecting TMH-derived
peptides in pathogens.
If this signature of selection is present in MHC-II is unknown.
This study shows that MHC-I [also has/does not have] more
epitopes derived from a TMH for a pathogen proteome, when compared with
a host proteome.
Additionally, MHC-II binds to peptides derived from TMHs 
[less/equally/more] often than expected by chance.
Lastly, we show the TMHs are [more/less/equally] 
conserved.
Our findings suggest that the immune system is [less/neutral/more]
vigilant to TMHs than expected by chance and this has 
left [a clear/a weak/no]
signal in the evolutionary history of the pathogen.

@@

\paragraph{Immune response}

Our immune system fights invaders on a daily basis.
These invaders can be fungi, bacteria or viruses.
The innate immune response is its first general 
and immediate strategy, where the acquired immune response
needs time to develop its specialized and more effective
combat forces.

\paragraph{Immune response by MHC-I}

All nucleated cells in humans present randomly sampled peptides
fragments to the surroundings of the cell using the Major 
Histocompatibility Complex (MHC) class I molecules.
If a cell gets infected by a virus, also the virus' peptides
will be presented at the cell surface. The foreign viral antigen is detected 
by cytotoxic T lymphocytes, which will kill the infected cell.

\paragraph{Immune response by MHC-II}

All pathogens can be detected by the foreign proteins on 
their (bacterial or fungal) cell walls or (viral) envelope.
For bacteria and fungi, this is the main mode of their detection,
as these do not infect a cell with their (foreign) DNA.
T and B cells express MHC-II to detect foreign peptide fragments.
An immune response is started when MHC-II detects a pathogen.

@@

Any human's immune system detects only a fraction of all possible
peptide fragments.
For humans, the MHC proteins are encoded by the
HLA ('Human Leukocyte Antigens') genes.
Each MHC complex can only bind a subset of all possible peptides.
For example, HLA-A and HLA-B have no overlap in which
peptides they bind (\cite{lund2004definition}).
The HLA region of humans is highly polymorphic, 
making it hard to classify all of the many haplotypes.
Classification of HLAs is based on algorithms that
maximize the information content of each 
classification, such as the presence of a certain amino acid at 
the first (for MHC-II, \cite{southwood1998several})
or second position (for MHC-I, \cite{lund2004definition}) of a peptide.
However, this does not imply that when a peptide follows
a classification pattern, that it will bind to the MHC. 
Instead, software should be used to predict if the
full peptide will do so.

>>>>

The human immune system is unexpectedly attentive towards
human transmembrane helices (TMHs), as TMH-derived peptide fragments
bind to histocompatibility complex (MHC) class I more often then expected
by chance only, which can be mostly explained by the low hydrophobicity
of those residues.
The physiological reason is yet unclear,
yet this finding hints that there has been selection upon detecting TMH-derived
peptides in pathogens.
If this signature of selection is present in MHC-II is unknown.
This study shows that MHC-I [also has/does not have] more
epitopes derived from a TMH for a pathogen proteome, when compared with
a host proteome.
Additionally, MHC-II binds to peptides derived from TMHs 
[less/equally/more] often than expected by chance.
Lastly, we show the TMHs are [more/less/equally] 
conserved.
Our findings suggest that the immune system is [less/neutral/more]
vigilant to TMHs than expected by chance and this has 
left [a clear/a weak/no]
signal in the evolutionary history of the pathogen.
The impact of this is [revolutionary/awesome].

@@

% \paragraph{Immune response}

Our immune system fights microbial pathogens on a daily basis,
such as fungi, bacteria or viruses.
The acquired immune response
needs time to develop its specialized and more effective
combat forces and learn to recognize the pathogens 
by use of the Major Histocompatibility Complex (MHC).

% \paragraph{Immune response by MHC-I}

Antigen-presenting cells such as dendritic cells
use the MHC class I molecules to 
present randomly sampled peptides fragments,
as created by the immunoproteasome (\cite{chapiro2006destructive}).
If a cell gets infected by an intracellular pathogen, also the virus' peptides
will also be presented at the cell surface. The foreign viral antigen is detected 
by cytotoxic T lymphocytes, which will kill the infected cell.

% \paragraph{Immune response by MHC-II, contrast with APC}

\richel{Weak, add refs}
All pathogens can be detected by the foreign proteins on 
their (bacterial or fungal) cell walls or (viral) envelope.
For bacteria and fungi, this is the main mode of their detection,
as these do not infect a cell with their (foreign) DNA.
T and B cells express MHC-II to detect foreign peptide fragments.
An immune response is started when MHC-II detects a pathogen.

@@

Any human's immune system detects only a fraction of all possible
peptide fragments.
For humans, the MHC proteins are encoded by the
HLA (Human Leukocyte Antigens) genes.
Each MHC complex can only bind a subset of all possible peptides.
For example, HLA-A and HLA-B have no overlap in which
peptides they bind (\cite{lund2004definition}).
The HLA region of humans is highly polymorphic, 
making it hard to classify all of the many haplotypes.
